\section{Introduction}
\label{sec:intro}

Network intrusion detection, botnet identification, and IoT malware analysis
are canonical cybersecurity tasks that operate over inherently
\emph{graph-structured} data: traffic flows between hosts form edges, devices
form nodes, and attack campaigns manifest as anomalous subgraph patterns.
Graph Neural Networks (GNNs) have emerged as powerful classifiers for
these tasks, achieving high accuracy on benchmarks such as CIC-IDS-2017
\citep{cicids2017}, UNSW-NB15 \citep{unswNB15}, and NF-BoT-IoT
\citep{nfbotiot} — yet they suffer from two critical, simultaneously
unresolved weaknesses.

\paragraph{W1 — Epistemic uncertainty from noisy and ambiguous features.}
Network telemetry is inherently imprecise: packet-level features exhibit
quantisation noise, flow statistics are aggregated over variable windows,
and labelling of attack vs.\ benign traffic is often soft or partially
overlapping \citep{graphids}.
Standard GNN classifiers produce crisp binary decisions that discard this
inherent fuzziness, leading to brittle predictions and inflated
false-alarm rates in production SOC deployments.

\paragraph{W2 — Absence of formal coverage guarantees.}
Neither standard GNNs nor fuzzy-neural extensions provide
\emph{distribution-free, finite-sample guarantees} on their prediction sets.
A security analyst receiving an alert has no principled bound on the
probability that the true attack class lies within the model's output —
making risk quantification impossible for compliance-sensitive
(e.g., GDPR, NIS2) environments \citep{cp_survey}.

\subsection{Research Gaps}

Significant progress has been made independently along two complementary
lines: (a) \emph{conformal prediction for GNNs}
\citep{cfgnn,daps,clarkson23,rrgnn} and (b) \emph{fuzzy GNN architectures}
\citep{flgnn,fgat,fgnn_survey}.
However, their explicit fusion remains an open problem.
Table~\ref{tab:gaps} positions FC-GNN against the five closest prior methods.

\begin{table}[t]
\caption{Research gap: no prior work simultaneously provides coverage
         guarantees, fuzzy aggregation, and cybersecurity validation.}
\label{tab:gaps}
\small
\begin{tabular}{lccc}
\toprule
Method & Coverage & Fuzzy & Cyb.\ Bench. \\
\midrule
CF-GNN \citep{cfgnn}        & \checkmark & $\times$ & $\times$ \\
RR-GNN \citep{rrgnn}        & \checkmark & $\times$ & $\times$ \\
FL-GNN \citep{flgnn}        & $\times$ & \checkmark & $\times$ \\
FGAT   \citep{fgat}         & $\times$ & \checkmark & $\times$ \\
GraphIDS \citep{graphids}   & $\times$ & $\times$ & \checkmark \\
\midrule
\textbf{FC-GNN (ours)}      & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Contributions}

This paper makes four contributions:

\begin{enumerate}
\item \textbf{FC-GNN architecture.}  A novel GNN in which each
  message-passing layer incorporates a fuzzy-rule aggregation module that
  outputs a fuzzy membership distribution over classes per node, rather than
  a crisp softmax logit (Section~\ref{sec:method}).

\item \textbf{Fuzzy Mondrian Conformal Prediction (FMCP).}  A principled
  extension of Graph-Structured Mondrian CP \citep{rrgnn} to fuzzy
  nonconformity scores defined via a Sugeno-type fuzzy integral, producing
  calibrated prediction sets with per-cluster conditional coverage guarantees
  (Section~\ref{sec:fmcp}).

\item \textbf{Coverage theorem.}  A finite-sample theorem proving that FC-GNN's
  prediction sets achieve at least $(1-\alpha)$ marginal coverage and
  conditional coverage within graph communities, with a slack term
  $\delta(|Q_m|) = O(|Q_m|^{-1/2})$ (Section~\ref{sec:theorem}).

\item \textbf{Cybersecurity benchmarks.}  Systematic evaluation on seven
  publicly available datasets spanning intrusion detection, botnet detection,
  and IoT malware, with temporal chronological splits and full ablation
  studies (Section~\ref{sec:experiments}).
\end{enumerate}

\subsection{Paper Structure}

Section~\ref{sec:related} reviews related work.
Section~\ref{sec:prelim} covers necessary preliminaries.
Section~\ref{sec:method} presents the FC-GNN architecture and the FMCP
calibration procedure.
Section~\ref{sec:experiments} reports empirical results.
Section~\ref{sec:interpretability} analyses the extracted fuzzy rules.
Section~\ref{sec:discussion} discusses limitations.
Section~\ref{sec:conclusion} concludes.
