\section{Related Work}
\label{sec:related}

\subsection{Conformal Prediction — Foundations}

Conformal prediction (CP) \citep{angelopoulos21} provides distribution-free,
finite-sample prediction sets satisfying marginal coverage
$\Prob(y \in \Calpha(x)) \ge 1-\alpha$ for any pre-specified error rate
$\alpha$.
\citet{barber23} extended CP beyond exchangeability to handle covariate shift.
\citet{cp_survey} survey CP comprehensively including a graph-data subsection.
The present authors have applied CP to intrusion detection \citep{dang23ids}
and botnet detection \citep{dang24botnet}, and developed CONFIDE for causal
competing-risk settings \citep{dang26confide}.

\subsection{Conformal Prediction for Graph Neural Networks}

\citet{cfgnn} introduced CF-GNN, the de-facto baseline for CP on graphs.
They derive a \emph{permutation-invariance condition} under which calibration
nonconformity scores are exchangeable and hence CP is valid; they also propose
a topology-aware score correction.
\citet{daps} proposed DAPS, which diffuses conformity scores over the graph
using homophily weighting, improving prediction-set tightness on node
classification.
\citet{clarkson23} studied structure-aware inductive node CP.
\citet{snaps} introduced SNAPS, aggregating scores from feature-similar
neighbours to improve singleton-hit proportion.

Among the closest to our work, \citet{rrgnn} proposed RR-GNN, which uses
\emph{Graph-Structured Mondrian CP} — partitioning calibration nodes into
graph communities and computing per-community quantiles.
This yields conditional coverage guarantees within communities and motivates
the FMCP component of FC-GNN.
\citet{rrgnn} do not, however, incorporate fuzzy aggregation or cybersecurity
evaluation.

Other CP-GNN work includes federated settings \citep{akgul25}, temporal GNNs
\citep{ncp2025,ncpnet25}, and conformal training \citep{confrank25}.
\citet{crcs_gad} extend CP to graph anomaly detection with risk control.
\citet{fuzzy_cp_evalue} develop fuzzy prediction sets via e-values in
non-graph settings, providing a theoretical bridge to our approach.

\subsection{Fuzzy Graph Neural Networks}

\citet{flgnn} proposed FL-GNN, which defines a Cartesian-product rule base
in the message-passing layer, enabling interpretable IF-THEN explanations.
Our fuzzy message-passing layer (FMPL) directly extends FL-GNN's rule base,
adding Gaussian membership functions, the product T-norm, and centroid
defuzzification.
\citet{fgat} extended fuzzy logic to graph attention (FGAT), using fuzzy
rough-set attention weights with dynamic negative sampling.
\citet{mfgat} proposed MFGAT with multi-view fuzzy attention.
A comprehensive survey by \citet{fgnn_survey} identifies the absence of
coverage guarantees as a key open problem across all fuzzy GNN architectures
— a gap we directly address.
\citet{firegnn} developed FireGNN with trainable neuro-symbolic fuzzy rules,
and \citet{fgenconv} applied explainable fuzzy GNNs to infrastructure
anomaly detection.
The present authors applied fuzzy embeddings to software-defined network NIDS
in \citep{dang24fuzzy}.

\subsection{GNNs for Cybersecurity}

\citet{graphids} achieved 99.98\% PR-AUC on NF-BoT-IoT using self-supervised
E-GraphSAGE.
\citet{gnni ds} proposed a hybrid static-dynamic graph IDS.
\citet{xgnid} combined heterogeneous GNN with LLM for dual-modality NIDS.
Various GNN-IDS architectures have been surveyed in \citep{gnn_ids_survey}.
The authors previously proposed GNN-based NIDS baselines in
\citep{dang23wifi} and federated IoT malware detection in
\citep{dang24federated}.

\subsection{Conformal Prediction in Cybersecurity}

\citet{botnet_cp} applied a CP rejection layer to botnet classification,
rejecting 58\% of incorrect predictions on ISCX.
\citet{cp_drift} showed that CP pseudo-labels become inconsistent under heavy
concept drift, motivating our temporal split protocol.
\citet{damba26} extended open-set CP to zero-day IIoT detection.
\citet{cp_cps26} proposed sliding-window temporal quantile CP for
cyber-physical system anomaly detection.

\subsection{Fuzzy Methods in Cybersecurity}

\citet{fuzzy_iot} combined fuzzy logic with CNN ensembles for IoT NIDS,
achieving 99.72\% on NSL-KDD.
\citet{fuzzy_sdn} proposed hunger-games-optimised fuzzy-rule deep learning
for SDN-enabled IoT environments.
\citet{gfs_gan} evolved fuzzy rules with adversarial training for IDS.
These works confirm the practical value of fuzzy reasoning for cybersecurity
but lack any CP coverage guarantees.
